\section{Evaluation}\label{sec:evaluation}
This section reports what the precompile-enabled zkVM regime can and cannot do for PLUM-based BDEC on consumer hardware, and uses the measurements to revisit the cost framework of Section~\ref{sec:theoretical}. Consistent with the rest of the thesis, the contribution is the \emph{quantification}: within the scope of the literature we surveyed, these are among the first reported figures for executing and proving a PLUM-instantiated post-quantum anonymous-credential workload inside a general-purpose zkVM, and we report them honestly, including the configurations that do not complete.

\subsection{Measurement Methodology}\label{ssec:eq}

The evaluation addresses four questions: whether PLUM verification, and the BDEC credential relation built on it, run inside the zkVM on the target hardware, with and without the precompile suite; where the zkVM cost falls and how much of it the precompile suite removes; for the configurations that do not complete, what the binding resource is and where the boundary lies; and how the two regimes compare on the deployment-lifecycle dimension of Definition~\ref{def:update-churn}.

\paragraph{Two measurement modes, and when each is reportable.}
A zkVM runs in two modes relevant here. \emph{Execute mode} runs the guest and
counts RISC-V \emph{cycles} without producing a proof; its cost is
machine-independent and terminates within the guest's memory space. \emph{Prove mode} produces the
cryptographic receipt and is the deployment-relevant cost, but is orders of
magnitude heavier and, as reported below, does not always complete on the target
hardware. Our rule is therefore: \emph{report prove wall-clock where proving
completes, and fall back to execute-mode cycle counts (machine-independent and
reproducible) where it does not.} Cycle counts and wall-clock times are never
compared across this boundary as if commensurable; each table states its mode.
Throughout this section, \emph{feasible} means that the prove completes on the
target machine within its $24$~GB memory envelope; whether a completed time is
acceptable for a given deployment is a separate judgment, treated by the
decision framework of Section~\ref{sec:discussion}.

\paragraph{What we deliberately do not claim.}
We do not report a single ``$N\times$ speedup'' headline. As
\S\ref{ssec:headline} makes explicit, the with/without-precompile contrast at the
prove level is an \emph{infeasible-to-feasible} transition rather than a ratio of
two finite times, and the finite reductions we do report are cycle-level
reductions at a stated security level and mode. Conflating these into one
multiplier would produce the kind of setting-dependent figure we avoid.

\subsection{Experimental Setup}\label{ssec:setup}

\begin{table}[h]
\centering
\small
\begin{tabular}{ll}
\toprule
Component & Specification \\
\midrule
Hardware & Apple M5~Pro, 18-core CPU, 20-core GPU, 24~GB unified memory, 1~TB SSD \\
OS & macOS 26.5 (Darwin 25.5.0), Apple Silicon (arm64) \\
Prover & CPU-bound; no CUDA/network prover (Apple Silicon has no NVIDIA GPU path) \\
zkVMs & SP1 v6.2.1 (forked), RISC~Zero v3.0.x \\
Signature & PLUM over $\mathbb{F}_p$, $p=2^{64}p_0+1$ ($199$-bit; see \S\ref{ssec:bigint}) \\
Security levels & $\lambda=80$ (prove-mode runs), $\lambda=128$ (execute-mode cycle attribution) \\
\bottomrule
\end{tabular}
\caption{Evaluation platform. All prove-mode measurements are local
single-machine runs with no network or dev-mode acceleration unless stated.}
\label{tab:platform}
\end{table}

\subsubsection{Settings Justification}\label{ssec:settings-justify}
A measurement is only persuasive if its settings are ones a deployer would
plausibly choose, rather than ones selected to flatter a number. We state and
justify each fixed parameter, and disclose where a setting is tuned.

\begin{description}
  \item[Consumer hardware (M5~Pro, 24~GB).] The research question concerns
  deployability on a personal machine, so this hardware is the thesis's explicit
  scope. Every feasibility claim is scoped to this envelope and is not asserted
  in general.
  \item[Security level: $\lambda=80$ for prove-mode, $\lambda=128$ for cycle
  attribution.] We disclose this seam rather than hide it. Prove-mode runs use
  $\lambda=80$ because $\lambda=128$ proving is untested and expected to exceed
  the memory budget, the precompile-free arm already failing at $\lambda=80$
  (\S\ref{ssec:frontier}); $\lambda=80$ is the largest level at which
  prove-mode wall-clock is observable on this hardware, which is why we use it
  and not because it is favourable. The cost-attribution analysis (\S\ref{ssec:attribution})
  is reported at the audience-relevant $\lambda=128$ because it is
  machine-independent (execute mode) and so remains measurable where proving is
  not. No number mixes the two levels silently; each table states $\lambda$.
  \item[Prover memory tuning.] The base prove-mode runs used a documented
  reduced-memory configuration. At SP1's default settings the Cell~2 prove is
  OOM-killed after about $3$~minutes on the $24$~GB machine and yields no
  figure at all; the tuning (\texttt{SHARD\_SIZE}$\,=2^{22}$, default $2^{24}$;
  \texttt{ELEMENT\_THRESHOLD}$\,=2^{26}$; \texttt{HEIGHT\_THRESHOLD}$\,=2^{20}$;
  \texttt{TRACE\_CHUNK\_SLOTS}$\,=2$; \texttt{RAYON\_NUM\_THREADS}$\,=8$, default
  $18$; \S\ref{ssec:repro}) trades more, smaller shards and fewer concurrent
  threads for fitting the $24$~GB envelope, holding peak memory near $81\%$.
  The headline timings are therefore tuned-configuration times: Cell~2
  completes in $32.53$~min and the SHA-3 control in $13.28$~min under identical
  tuning. The \emph{zero-knowledge wrap} attempts ran under the same class of
  reduced configuration and still exhausted memory
  (\S\ref{ssec:resource-frontier}). A contributing factor to the Griffin arm's
  memory pressure surfaced during cost-model calibration: the deployed fork's
  shard-budgeting cost table prices the Griffin chip at $348$ trace columns,
  whereas the chip's compiled width is $3{,}819$, so the executor budgets a
  Griffin-bearing shard at roughly one tenth of its real committed area and
  splits shards later than the trace-area threshold intends. Correcting the
  table entry is a one-line change whose effect on Cell~2's peak memory and
  wall-clock is untested and left to future work.
  \item[Micro-benchmark isolation.] Cost-attribution runs isolate single
  primitives (one Griffin permutation; one verification) so that per-operation
  cost can be attributed rather than inferred from an aggregate.
\end{description}

\subsubsection{Reproducibility}\label{ssec:repro}
Prove-mode and attribution runs use fixed RNG seeds and fixed messages; the
SP1/RISC~Zero submodule versions are pinned in the lockfile; the memory-tuning
environment variables, host binaries, and the adversarial-probe harness are
recorded in the project's measurement notes, together with the per-run commit
hashes and, where applicable, guest-image identifiers.

\subsection{PLUM Verification With and Without the Precompile}\label{ssec:headline}

Table~\ref{tab:plum-verify} reports prove-mode results for a single PLUM
signature verification at $\lambda=80$. The central comparison is between the
precompile-free baseline and the precompile-enabled arm on the same hardware.

\begin{table}[h]
\centering
\small
\begin{tabular}{llllll}
\toprule
zkVM / mechanism & $\lambda$ & Prove time & Verify time & Peak mem & Outcome \\
\midrule
SP1, rv32im Griffin (no precompile)          & 80 & n/a       & n/a    & 87.94\% & DNF (OOM $\approx$1m45s) \\ % 582ad1a Cell 1
SP1, Griffin-$\mathbb{F}_p^{192}$ precompile & 80 & 32.53~min\textsuperscript{\dag} & 3.06~s & 81.31\% & receipt verified \\          % 582ad1a Cell 2
SP1, SHA-3 hasher (control)                  & 80 & 13.28~min & 2.80~s & 80.92\% & receipt verified \\ % 91941ed
RISC~Zero, \texttt{sys\_bigint}, SHA-3 hasher & 80 & 6\,h\,19\,m & n/a\textsuperscript{\ddag} & n/a   & receipt verified \\          % uint256 doc
\bottomrule
\end{tabular}
\caption{PLUM single-signature verification, prove mode, $\lambda=80$, M5~Pro
24~GB. We refer to the three SP1 arms as \emph{Cell~1} (no precompile),
\emph{Cell~2} (Griffin-$\mathbb{F}_p^{192}$ precompile), and \emph{Cell~3}
(software SHA-3 control) throughout. The precompile-free baseline does not complete; the precompile arm completes
in $32.53$~min. The SHA-3 arm is a \emph{control} (PLUM with a standard hash
computed in software, no hash precompile, only the shared \texttt{UINT256\_MUL}
field-arithmetic precompile active), not a deployable PLUM variant. These are
the May calibrated-reference runs; a $2026$-06-13 current-build re-measurement
gives Cell~2 $17.81$~min and Cell~3 $16.15$~min ($n=2$ per arm, same session),
confirming feasibility and the inversion sign, the difference attributed to
build drift (\S\ref{ssec:headline}). $^\dag$A later clean-machine
re-measurement on 2026-06-18 completed the same Cell~2 prove in $14.89$~min
($18.67$~GB peak RSS, verify OK); the Cell~2 prove time is accordingly reported
elsewhere in this thesis as the $14.89$--$32.5$-min range, with the $32.53$-min May run
retained as the calibrated reference. $^\ddag$RISC~Zero receipt-verify wall-clock not separately logged; the
receipt verified.}
\label{tab:plum-verify}
\end{table}

A re-run of the precompile-free arm on 2026-06-12, under the documented
reduced-memory tuning (\S\ref{ssec:settings-justify}), ran for
$5$\,h\,$07$\,m without terminating and without exhausting memory before we
stopped it by hand, yielding no receipt (run record in
\texttt{docs/measurements/audit5\_campaign\_20260612}). Under that tuning the
binding resource is time rather than memory, and the prove still does not
complete within five hours on this hardware.

\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  width=0.85\textwidth, height=6.2cm,
  ymode=log, ymin=1, ymax=1000,
  ylabel={Prove time (min, log scale)},
  symbolic x coords={SHA-3 control, Griffin precompile, RISC Zero bigint (SHA-3)},
  xtick=data,
  x tick label style={font=\footnotesize},
  nodes near coords, point meta=explicit symbolic,
  every node near coord/.append style={font=\footnotesize},
  bar width=30pt, enlarge x limits=0.3,
]
\addplot+[draw=black, fill=gray!35] coordinates {
  (SHA-3 control, 13.28) [13.3 min]
  (Griffin precompile, 32.53) [32.5 min]
  (RISC Zero bigint (SHA-3), 379) [6\,h\,19\,m]
};
\end{axis}
\end{tikzpicture}
\caption{Prove time on the configurations that complete ($\lambda=80$, M5~Pro
24~GB, log scale). The bespoke Griffin precompile ($32.5$~min) is slower than the
SHA-3 control ($13.3$~min) at matched security, and RISC~Zero's \texttt{sys\_bigint}
route (also SHA-3-hashed) is slower again ($6$\,h\,$19$\,m); these are different hash and substrate
mechanisms, so the bars are not one artefact. The precompile-free baseline does not
complete (DNF, OOM at ${\approx}1$m$45$s) and so has no bar.}
\label{fig:prove-time}
\end{figure}

\paragraph{The honest reading.}
The with/without-precompile contrast is not a ratio of two finite times: without
the precompile the workload \emph{does not complete} on this hardware, and with
it the workload completes in $32.53$~minutes. The defensible claim is therefore
qualitative at the prove level (\emph{the precompile moves PLUM verification
from infeasible to feasible on consumer hardware}) and quantitative only at the
cycle level (\S\ref{ssec:attribution}). The two zkVMs are not directly comparable
here: the RISC~Zero leg is the SHA-3-hasher arm, carrying the field arithmetic
through \texttt{sys\_bigint} without a native Griffin chip
(\S\ref{ssec:griffin-air}), so its like-for-like SP1 counterpart is the
$13.28$-min SHA-3 control, with the residual same-hash gap reflecting the
substrate and receipt kind; no Griffin-hashed single-verification prove was run
on RISC~Zero (the Griffin AIR port is future work), and the gaps should not be
read as a like-for-like substrate ranking. The SHA-3 control
isolates the hash family: the verification harness is identical except that every
Griffin-$\mathbb{F}_p^{192}$ sponge call is replaced by a SHA3-256 call computed in
software, so that the SHA-3 arm fires \emph{no} hash precompile and only the shared
\texttt{UINT256\_MUL} field-multiplication precompile remains active. It
shows that a workload using a \emph{standard} hash in software is markedly
cheaper than one using the algebraic hash PLUM requires through a bespoke chip,
which is the quantitative form of the arithmetic-mismatch argument of Section~\ref{sec:theoretical}. The
inversion is direct: the bespoke Griffin-$\mathbb{F}_p^{192}$ precompile proves
\emph{slower} at prove than the software SHA-3 control at matched $\lambda=80$ in
every run we took, on both the May and the June builds. Its sign is robust; its
magnitude is build-sensitive, from $2.45\times$ across the May builds ($32.53$
versus $13.28$~min) to $1.10\times$ in a same-session June re-measurement
($17.81$ versus $16.15$~min, $n=2$ per arm, tight: the two Cell~2 runs at
$17.80$ and $17.82$~min, run records
\texttt{docs/measurements/audit5\_campaign\_20260612}). We attribute the
May-to-June difference on each arm to build and toolchain drift over the
interval, which we did not isolate; the $32.53$/$13.28$ figures remain the
calibrated reference of Table~\ref{tab:plum-verify} and the cost model below.
A clean-machine re-measurement on 2026-06-18 completed the same Cell~2 prove in
$14.89$~min at $18.67$~GB peak RSS (verify OK), $2.2\times$ faster than the
$32.53$-min May run; the gap is build- and machine-state-sensitive and we did not
isolate it, so the prove time is reported as the $14.89$--$32.53$-min range.
The inversion is robust in sign across runs while its magnitude ($1.10$--$2.45\times$) varies with the build, so it is reported as an envelope rather than a single factor.
We read the inversion as a genuine
property of the pipeline rather than a measurement artefact, and the trace-area
accounting of Section~\ref{sec:theoretical}
(Equation~\eqref{eq:trace-area-model}) postdicts its sign from the chips'
committed dimensions. The Griffin chip emits $14$ rows of $3{,}819$ columns per
permutation (each row carries $26$ modular-arithmetic gadgets of $127$ cells
each, beside range checks and state limbs), so the $1{,}052$ permutations of one
verification commit $14\times 3{,}819\times 1{,}052\approx 5.6\times10^{7}$
cells, plus one controller row per permutation, that the SHA-3 arm never pays.
The shared \texttt{UINT256\_MUL} chip is common-mode ($69{,}396$ versus
$69{,}385$ events) and cancels, and the Griffin arm additionally executes $13\%$
more guest cycles ($1.26\times10^{8}$ versus $1.11\times10^{8}$, execute mode).
Both terms of Equation~\eqref{eq:trace-area-model} are therefore larger on the
Griffin arm, so the model places it behind the control for every non-negative
choice of the per-cell rates, granted a common per-cycle core area across the
two arms; the calibration of the model's constants on these two runs is
reported with the cost model (\S\ref{ssec:costmodel}). The feasibility recovery and the inversion are
the same inequality (Proposition~\ref{prop:precompile-pays}, Equation~\eqref{eq:precompile-pays}) evaluated at two
operating points. The precompile bought a finite measurement, and the
accounting prices why a speedup over the control was never available: against a
control whose hash costs the substrate no chip area at all, the Griffin arm
pays the chip area and recovers nothing.

\paragraph{Proof size, and why we do not re-measure it.}
Receipt size in SP1 is a property of the proof \emph{type}, not of the workload.
The runs above use SP1's default \emph{core} proof, whose size \emph{scales with
execution size}~\cite{succinct2025prooftypes}; the compact, on-chain-deployable
proof is instead the constant-size {SNARK} \emph{wrap} of that receipt, about
$260$~bytes for Groth16 over {BN254}, or $868$~bytes for
{PLONK}~\cite{succinct2025prooftypes}. Producing that wrap is exactly the step
that does not complete on the target hardware as the toolchain ships. The
failure is not a want of memory; it is that SP1's recursion shapes are
provisioned for smaller relations than PLUM verification, the
recursion-shape provisioning wall
of \S\ref{ssec:resource-frontier}. On a
$24$~GB machine, then, the proof a deployer can actually generate is the large,
shard-scaling core receipt; the $260$-byte {SNARK} is a documented size the
frontier of \S\ref{ssec:resource-frontier} prevents us from realising here. We therefore report the deployable figure as
documented and do not serialise the workload-dependent core receipt, whose byte
count is not the deployment-relevant number.

\subsection{The Circuit-SNARK Reference Point}\label{ssec:aurora-ref}

To situate the zkVM costs against the static-circuit regime they are meant to
escape, we measured the Aurora prover end-to-end on the same hardware over a
single Loquat-verification R1CS. This is the single-verification building block
the BDEC construction composes~\cite{10.1007/978-981-96-0957-4_3}: the full
credential relation merges two such verifications under a shared hidden public
key, so it is roughly twice this figure, and we did not run the composed relation
statically. We measure it directly rather than cite it, so that the comparison is
on our hardware rather than across machines.

\begin{table}[h]
\centering
\small
\begin{tabular}{lr}
\toprule
Quantity & Value \\
\midrule
Constraints (padded)        & $524{,}288$ \\
Prover time (setup $+$ prove) & $225.4$~s ($3.76$~min; $223.2$--$226.6$) \\
Verifier time               & $41.3$~s ($41.0$--$41.6$) \\
Proof size                  & $704{,}416$~B ($688$~KiB, identical across runs) \\
Achieved security           & $107.5$ bits (target $128$) \\
Zero-knowledge              & disabled \\
\bottomrule
\end{tabular}
\caption{Aurora over a single Loquat-verification R1CS, M5~Pro 24~GB; mean of
three solo runs with the per-run range in parentheses (run records in
\texttt{docs/measurements/pub\_hardening\_20260610}).
This is the \emph{Loquat}-instantiated circuit over $\mathbb{F}_{p^2}$ ($127$-bit
base)~\cite{10.1007/978-981-96-0957-4_3} at single-verification scope; the full
BDEC relation is two such verifications ($\approx$twice this), not run statically.
Prover time includes Aurora parameter setup but is not a recompile-time
measurement (\S\ref{ssec:churn-eval}). Achieved security $107.5$ bits, below the
$128$ target. Verification \texttt{ACCEPT} in every run.}
\label{tab:aurora-ref}
\end{table}

\paragraph{The zero-knowledge-enabled run.}
The runs of Table~\ref{tab:aurora-ref} have zero knowledge disabled. We
additionally ran the same $2^{19}$ R1CS ($524{,}288$ constraints,
$\mathsf{RS}$ extra dimensions $5$) with zero knowledge \emph{enabled}, three
solo runs: the prover takes $1317.7$~s at the mean ($21.96$~min;
$1285.7$--$1376.1$~s), the verifier $40.8$--$41.8$~s, and the proof is
$876{,}768$~B ($856$~KiB), identical across the three runs, against
$704{,}416$~B with zk disabled; achieved security is $107.5$ bits (target
$128$, as in the zk-disabled legs), peak RSS spans $9.2$--$11.0$~GiB,
comfortably inside the $24$~GB envelope, and the verifier returns
\texttt{ACCEPT} in every run (run records in
\texttt{docs/measurements/pub\_hardening\_20260612} and
\texttt{audit5\_campaign\_20260612}). Against the zk-disabled
prover mean of $225.4$~s, the measured zero-knowledge premium on this circuit
is about $5.8\times$ prover time at the means.

\paragraph{A same-scheme substrate comparison.}
To remove the scheme confound, we ran the \emph{same} Loquat verification through
the zkVM, at the same parameters. The result is a clean single-axis comparison
(Table~\ref{tab:loquat-substrate}): for one Loquat verification at $\lambda=80$,
with neither side zero-knowledge, the static Aurora circuit proves in $3.76$~min
against $44$--$64$~min for the SP1 zkVM ($51.9$~min at the mean of three runs;
the spread on an otherwise idle machine is thermal, run~2 starting on a package
heated by run~1). Same scheme, same $\mathbb{F}_{p^2}$
($127$-bit) field, same unit, same security setting; only the substrate differs.
The zkVM is the dearer per-proof substrate by $12$--$17\times$
($13.8\times$ at the means), the gap being
the cost of executing Loquat's field arithmetic as emulated RISC-V over KoalaBear
rather than as a native constraint system. The zkVM leg uses no Loquat-specific
precompile (none exists for the $127$-bit field), so this is the substrate's
emulated cost, which is precisely what adopting a general-purpose zkVM both buys
and pays for.
The two substrates are measured at different rates, the zkVM arm at $|U|=4096$ and the Aurora arm at $|U|=256$, so the ratio indicates direction rather than a calibrated factor.
The instance uses the paper-faithful rate after the $\rho^{*}$ fix but keeps $L$ and $B$ reduced below the paper's $\lambda=128$ set, so the timing is not fully security-calibrated.

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Substrate & Prove time & Proof size & ZK \\
\midrule
Aurora static circuit (native $\mathbb{F}_{p^2}$) & $3.76$~min & $688$~KiB & no \\
SP1 zkVM ($\mathbb{F}_{p^2}$ emulated on KoalaBear) & $44$--$64$~min & $405$~MiB & no \\
\bottomrule
\end{tabular}
\caption{Same-scheme substrate comparison: one Loquat verification, $\lambda=80$,
M5~Pro 24~GB, three solo runs per leg, neither zero-knowledge. The Aurora leg is
the mean of Table~\ref{tab:aurora-ref}; the SP1 leg is the range of three runs
($44.4$, $47.4$, $64.0$~min) under the documented reduced-memory configuration
(\S\ref{ssec:settings-justify}), succinct core STARK, peak RSS
$14.1$--$15.4$~GiB, execute-mode cost $397{,}227{,}544$ cycles (run records in
\texttt{docs/measurements/pub\_hardening\_20260610}). Only the substrate differs
(native R1CS versus zkVM emulation); the $12$--$17\times$ prover-time gap is the
cost of the zkVM substrate for this scheme. This isolates the substrate for
\emph{Loquat}; it does not transfer to PLUM, which has no $127$-bit Aurora harness.}
\label{tab:loquat-substrate}
\end{table}

\paragraph{What this reference does and does not license.}
Two readings of this point are legitimate and one is not. \emph{Legitimate:} on this
hardware the static-circuit regime proves a single Loquat verification in
minutes, against the zkVM's tens-of-minutes-to-infeasible
(Tables~\ref{tab:plum-verify},~\ref{tab:bdec}), so for a \emph{fixed} predicate
the static circuit is the faster prover, as one would expect from a specialised
encoding. \emph{Also legitimate:} the zk-disabled instance of
Table~\ref{tab:aurora-ref}, like the zkVM succinct receipt, does not by itself
discharge BDEC's anonymity requirement; the zero-knowledge-enabled run reported
above supplies that property for the same circuit, at the measured premium. \emph{Not licensed:} reading
Table~\ref{tab:aurora-ref} \emph{against the PLUM zkVM numbers}
(Tables~\ref{tab:plum-verify},~\ref{tab:bdec}) as a same-scheme verdict, since that
pairs Loquat over a $127$-bit field with PLUM over a $199$-bit field. The
same-scheme verdict is instead the Loquat-versus-Loquat comparison of
Table~\ref{tab:loquat-substrate}; either reading is timed without the
predicate-change cost that is the separate subject of the update-churn comparison
(\S\ref{ssec:churn-eval}).

\subsection{Cost Attribution}\label{ssec:attribution}

To attribute the cost we use execute-mode cycle counts at $\lambda=128$, the
audience-relevant security level, where prove-mode is infeasible but cycle counts
remain machine-independent. Table~\ref{tab:ablation} reports the cumulative
effect of routing PLUM's field arithmetic through the host precompile, as an
ablation over the SHA-3-hasher verification path.

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Configuration (cumulative) & SP1 cycles & RISC~Zero cycles & $\Delta$ vs.\ baseline \\
\midrule
Baseline (\texttt{BigUint} emulation)        & 289{,}778{,}709 & 812{,}646{,}400 & n/a \\          % uint256 doc
$+$ field mul via host precompile            & 276{,}643{,}994 & n/a             & SP1 $-4.5\%$ \\
$+$ power as square-and-multiply             & 240{,}857{,}140 & n/a             & SP1 $-16.9\%$ \\
$+$ skip redundant reduction                 & 179{,}952{,}920 & 239{,}599{,}616 & SP1 $-37.9\%$, R0 $-70.5\%$ \\
\bottomrule
\end{tabular}
\caption{Cycle-level cost attribution, PLUM verification, SHA-3 hasher,
$\lambda=128$, \emph{execute mode}. Routing field arithmetic through the host
precompile (and removing redundant reductions) cuts $37.9\%$ of SP1 cycles and
$70.5\%$ of RISC~Zero cycles; RISC~Zero's larger reduction reflects its heavier
per-multiplication baseline emulation ($6{,}625$ cycles/mul at baseline, falling
to $1{,}826$ with the precompile path, against SP1's baseline of about
$2{,}567$ cycles/mul).}
\label{tab:ablation}
\end{table}

\paragraph{Where the cost lives, and the parameter-set caveat.}
At $\lambda=128$ a single PLUM verification on the SHA-3-hasher arm issues
about $112{,}902$ $\mathbb{F}_p^{192}$ multiplication syscalls through the host
precompile, as counted in our own SP1 run log; no Griffin permutation executes
on that arm, so this figure measures the verifier's \emph{non-hash} field
arithmetic (PLUM \S4.2 reports $\approx 10^4$ non-hash algebraic R1CS
constraints, $10{,}333$ of $116{,}285$). The Griffin-internal multiplication
cost appears instead in the with/without-AIR contrast at $\lambda=80$: the
software-Griffin arm fires $4{,}870{,}724$ \texttt{UINT256\_MUL} syscalls
against $69{,}396$ with the Griffin AIR active, the hash-internal field
operations either absorbed by the \texttt{GRIFFIN\_FP192\_PERMUTE} syscall or
inflating the multiplication count $70\times$. The contrast is therefore
consistent with the arithmetic-mismatch model: the cost is dominated by the
algebraic hash, whose internal field operations leave the syscall stream only
when the bespoke chip absorbs them. We caution explicitly against transferring the PLUM paper's
circuit-SNARK figure (that $91\%$ of \emph{R1CS constraints} are hash
work) directly to the zkVM cost model: that figure is a constraint count in a
different proving system, whereas Table~\ref{tab:ablation} is a RISC-V cycle
count. The qualitative conclusion (hash/field arithmetic dominates) agrees; the
exact share differs by cost model and is reported here in the zkVM's own units.

\paragraph{Execute-mode delta for the Griffin precompile (PLUM-80).}
For the bespoke Griffin AIR specifically, the precompile-free versus
precompile-enabled execute-mode cycle counts differ by a factor of
$56.4\times$ at $\lambda=80$, the ratio of precompile-free to Griffin-AIR-enabled cycles on the Griffin arm, consistent with the field-mismatch lower bound of Proposition~\ref{prop:fmt-lb} (which predicts a per-multiplication tax of up to ${\approx}49$ for PLUM's $199$-bit field on BabyBear); this is distinct from the $37.9\%/70.5\%$ of Table~\ref{tab:ablation}, a separate field-multiplication-reduction ablation reported at $\lambda=128$. % commit 2c1392b
We report this as an \emph{execute-mode cycle ratio} rather than a wall-clock speedup:
it is the machine-independent counterpart of the infeasible-to-feasible prove
transition of Table~\ref{tab:plum-verify}, and it is the figure that should be
cited when a single number is wanted, in place of any prove-time ``speedup,''
which is undefined when the baseline does not complete.

\subsection{PLUM-in-BDEC and the Consumer-Hardware Frontier}\label{ssec:frontier}

Moving from a single signature to the BDEC credential-generation relation
$R_{\mathsf{cre}}^{\mathsf{PLUM}}$ (two signature verifications under a hidden
user key; \S\ref{ssec:integration}), Table~\ref{tab:bdec} reports the
application-level results at $\lambda=80$: RISC~Zero in execute mode and SP1 in
succinct prove mode for both BDEC relations.

\begin{table}[h]
\centering
\small
\begin{tabular}{l l p{3.2cm} p{4.6cm}}
\toprule
Workload & Mode & Cost & Outcome \\
\midrule
BDEC CreGen, RISC~Zero & execute & $9.40\times10^{9}$ cyc, $\approx$65~s & both verifications accept \\ % plum_in_bdec doc
BDEC CreGen, RISC~Zero & prove (succinct) & $\geq$15.6~h, peak RSS 11.24~GB & internal-verifier rejection (anomaly; \S\ref{ssec:cregen-anomaly}) \\ % risc0_prove_failure doc
BDEC CreGen, SP1 & prove (succinct) & $26.98$ and $31.11$~min ($n=2$) & receipt verified; not zero-knowledge \\ % S9 pub_hardening_20260611 + audit5
BDEC ShowCre $k{=}1$, SP1 & prove (succinct) & $40.15$ and $44.24$~min ($n=2$) & receipt verified; not zero-knowledge \\
BDEC ShowCre $k{=}2$, SP1 & prove (succinct) & $58.14$~min ($n{=}1$) & receipt verified; not zero-knowledge \\
\bottomrule
\end{tabular}
\caption{BDEC credential relations, $\lambda=80$, M5~Pro 24~GB. On RISC~Zero,
execute mode validates the honest-input path of CreGen ($9.4\times10^{9}$
cycles, ${\approx}4.69\times10^{9}$ per signature verification); its succinct
prove yielded no receipt, from an internal-verifier rejection rather than a
resource limit (\S\ref{ssec:cregen-anomaly}), not used as a feasibility result.
On SP1 both relations complete a \emph{succinct} prove: CreGen in $26.98$ and
$31.11$~min, ShowCre in $40.15$ and $44.24$~min at $k{=}1$ and $58.14$~min at
$k{=}2$. These receipts are succinct STARKs and are \emph{not} zero-knowledge;
the $k{=}2$ figure agrees with the additive cost model within ${\approx}3.3\%$
of the $(k{+}2)/3$ projection from the $k{=}1$ mean (run records
\texttt{docs/measurements/pub\_hardening\_20260611} and
\texttt{audit5\_campaign\_20260612}).}
\label{tab:bdec}
\end{table}

The execute-mode result validates the honest path of the BDEC credential
relation end-to-end inside the zkVM: both signature verifications accept, at
$9.4\times10^{9}$ cycles; rejection behaviour was probed only at
single-verification scope (Section~\ref{sec:system}). The prove-mode outcome requires care,
because it is \emph{not} the deployability frontier it might first appear to be, and
we separate the two readings explicitly.

\subsubsection{An internal-verifier rejection rather than a deployability frontier}\label{ssec:cregen-anomaly}
The succinct-prove run (RISC~Zero $3.0.5$, the resolved version of the repository lockfile, verified against the compiled library; \texttt{ProverOpts::succinct()}, the
succinct-STARK path; RISC~Zero's API default is the composite receipt)
terminated after $15.6$~hours with
\texttt{verify segment / verification indicates proof is invalid}. The error
context localises the failure precisely: the string \texttt{verify segment} is
attached at a single site in the prover
(\texttt{prover\_impl.rs:280}, identical in releases $3.0.3$ through $3.0.5$),
the integrity self-check a freshly proved \emph{base segment} must pass
immediately after \texttt{prove\_core}; the recursion layers carry distinct
contexts (\texttt{verify lift}, \texttt{verify join}) that never appeared. The
prover therefore rejected one ordinary segment of its own making, well
mid-stream: at the measured ${\approx}40$~s per segment, $15.6$~hours
corresponds to roughly segment $1{,}400$ of $9{,}936$. This is
categorically different from running out of memory: peak resident memory was only
$11.24$~GB on the $24$~GB machine, with kernel free memory at $70$--$85\%$ for
nearly the whole run and dipping to $47\%$ only during a brief
memory-compression spike in the first minutes, so the machine was \emph{not}
the limiting resource. An
internal-verifier rejection is a correctness anomaly; it tells us nothing about
deployability, and we therefore do \emph{not} report it as a frontier or fold it
into any feasibility claim. The localisation also rules candidates out. A
segment-count threshold is contradicted by the index arithmetic: the rejection
occurred near segment $1{,}400$, below the ${\approx}4{,}500$ segments the
single-verification workload clears (the RISC~Zero row of
Table~\ref{tab:plum-verify}, $6$\,h\,$19$\,m, receipt verified). A
content-deterministic cause, including a malformation of the
\texttt{sys\_bigint} precompile trace (the bespoke Griffin AIR named by
Assumption~\ref{asm:air} of Section~\ref{sec:security} is SP1-only and was not on this
proving path), is disfavoured by the partial composite re-run (started
2026-05-31): with the same guest, the same deterministically seeded input, and
the same machine, it proved $2{,}452$ of the $9{,}936$ segments
($\approx$$24.6\%$, covering the region where the succinct run failed) through
the same per-segment integrity check, in $34.7$~h at a steady ${\approx}40$~s
per segment, peak RSS ${\approx}9.9$~GB, without reproducing the rejection,
before we terminated it (projection ${\approx}4.6$~days for all segments). The
surviving candidates are therefore (i) a transient error over the
${\approx}245$ CPU-hour run on non-ECC memory, and (ii) a concurrency-sensitive
defect in the succinct path's overlapped prove/lift/join pipeline, the one
structural element the composite path does not exercise. The failing-segment
index was not captured in the succinct run; the discriminating next step is a
succinct re-run with segment-index logging and the recursion stages serialised,
which separates (i) from (ii) by whether the rejection recurs and where.
Pending it, the anomaly is recorded as exactly that (an anomaly under
diagnosis), and the thesis's claims do not rest on it.

\subsubsection{The genuine deployability frontier}\label{ssec:resource-frontier}
Two failures bound the deployability frontier of this thesis, both unambiguous
as to cause and independent of the anomaly above. The first is the PLONK
zero-knowledge wrap of the standalone PLUM single-signature-verification receipt
(Cell~2 of Table~\ref{tab:plum-verify}, the $32.53$-min prove, not CreGen or
ShowCre). At SP1's default prover concurrency this wrap is killed under memory
pressure at ${\approx}83\%$ of the $24$~GB budget (peak RSS $19.98$~GB per the
process accounting at the kill; a periodic RSS sampler observed ${\approx}17.8$~GB
before the final spike; run records
\texttt{docs/measurements/bench\_suite\_20260529\_overnight} (sampler) and
\texttt{docs/measurements/zkwrap\_worker\_concurrency\_20260627}). This kill is
not, however, a memory limit of the workload: lowering the recursion-worker count
runs the same wrap under ${\approx}17$~GB, comfortably inside the budget
(\texttt{docs/measurements/zkwrap\_default\_shard\_20260627}), where it fails earlier
instead, at the recursive-compression step, because SP1's static recursion
shapes are provisioned for smaller relations than PLUM verification; we call this
\textbf{the recursion-shape provisioning wall}. Bringing the
wrap within reach is a one-time regeneration of those static artifacts (the
recursion-shape and verifying-key tables), roughly $51$~hours of compute on this
hardware ($48$--$54$~h, from a measured per-setup rate of ${\approx}1$~s in a
single CPU process over the $185{,}862$-entry rebuild; the measured rate is in run
record \texttt{docs/measurements/zkwrap\_vkmap\_slope\_20260628} and the extrapolation
to $\approx 51$~h in \texttt{docs/measurements/zkwrap\_failure\_map\_20260628}), rather than more memory. We do not run the regenerated PLUM toolchain end-to-end, but the wrap's two memory-heaviest stages each fit the budget. PLUM's core-and-recursion proving peaks at ${\approx}17$~GB, and the program-independent gnark {PLONK} wrap stage at $15.4$~GB peak {RSS} (a trivial-guest probe of the fixed BN254 wrap circuit, $n{=}1$; \texttt{docs/measurements/gnark\_plonk\_wrap\_ram\_20260629}). No stage exceeds $24$~GB, so the binding barrier is the one-time ${\approx}51$~h regeneration, a compute cost that more memory does not remove. The
provisioning wall specific to the anonymity-providing wrap is thus structural, not a memory
ceiling, and it is the measured form of the anonymity--deployability tension of
Section~\ref{sec:security}.

\paragraph{The precompile is the source of the provisioning wall.}
The provisioning wall attributes entirely to the
Griffin-$\mathbb{F}_p^{192}$ precompile. Removing only the Griffin chip, with
every other fork edit intact, reproduces the pre-Griffin compose shape
byte-for-byte ($795{,}264$ provisioned ExtAlu events, md5 \texttt{f97a8b2f});
with the chip the shape is $1{,}072{,}032$, a marginal $+276{,}768$ ExtAlu
($+34.8\%$) and $+38{,}816$ PrefixSumChecks, while the fork's non-Griffin
recursion edits move the shape by exactly zero. The required height
($1{,}064{,}022$, an overflow of $268{,}758$ over the committed $795{,}264$
budget) is what trips the arity-$4$ compose shape against that budget (run
record \texttt{docs/measurements/zkwrap\_griffin\_shape\_attribution\_20260628}).
The precompile that lowers the core-layer cost is therefore the same chip that
enlarges the recursion shape the wrap must verify.

Second, and
separately, a \emph{non-zero-knowledge} baseline proof of a standardised-lattice
signature verification (an ML-DSA proxy: the $17$ NTT-domain polynomial
multiplications of $256$ coefficients modulo $q=8{,}380{,}417$ that dominate
ML-DSA-65 verification, proved as a baseline succinct STARK on SP1, not a
zero-knowledge wrap) was jetsam-killed after $249$~s with peak RSS $12.59$~GB
and a virtual-memory footprint of $82.7$~GB (compressed memory plus swap, hence
exceeding physical RAM);
this is a substrate base-cost datum that \emph{strengthens but does not
instantiate} the tension, since it bears on the cost of proving any PQ
verification on this substrate, not on the zero-knowledge wrap specifically. These are the data points the frontier claim uses;
the positive feasibility anchor remains the PLUM single-verification prove of
Table~\ref{tab:plum-verify}, which completes in $32.53$~min. Table~\ref{tab:proof-mode}
summarises, for every proof mode attempted, whether a receipt is produced and whether
it is zero-knowledge and post-quantum.

\begin{table}[h]
\centering
\small
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{4.2cm}p{1.4cm}p{1.6cm}p{1.5cm}p{2.9cm}@{}}
\toprule
Proof mode (substrate / configuration) & Produced? & Zero-know\-ledge? & Post-quantum? & Resource outcome \\
\midrule
SP1, PLUM verify, Griffin precompile (\texttt{core}) & yes & no & yes & completes, $32.5$~min \\
SP1, PLUM verify, SHA-3 control (\texttt{core}) & yes & no & yes & completes, $13.3$~min \\
SP1, PLUM verify, Groth16/PLONK wrap & \textbf{no} & yes & \textbf{no} & structural: recursion-shape provisioning wall (OOM ${\approx}20$~GB, default conc.) \\
RISC~Zero, PLUM verify, \texttt{sys\_bigint}, SHA-3 hasher (succinct) & yes & blinded; ZK not formally established\textsuperscript{$\ddagger$} & yes & completes, $6$\,h\,$19$\,m \\
RISC~Zero, BDEC CreGen (succinct) & \textbf{no} & n/a & n/a & $\geq15.6$~h, internal-verifier rejection \\
SP1, BDEC CreGen (succinct) & yes & no & yes & completes, $27$--$31$~min \\
SP1, BDEC ShowCre $k{=}1,2$ (succinct) & yes & no & yes & completes, $40$--$58$~min \\
Aurora, Loquat single-verify & yes & no\textsuperscript{$\dagger$} & yes & completes, $3.76$~min \\
Aurora, Loquat single-verify, ZK enabled & yes & yes & yes & completes, ${\approx}22$~min ($n{=}3$) \\
\bottomrule
\end{tabular}
\caption{Proof-mode feasibility matrix. Of the modes \emph{produced} on the target
hardware by the zkVMs, none is dependably zero-knowledge: the SP1 rows are not
zero-knowledge, the RISC~Zero
receipt carries no formally established zero-knowledge guarantee, and the only
zero-knowledge mode the zkVM toolchain exposes
(row~3, a pairing wrap) is neither post-quantum nor produced as the toolchain
ships, which is the dual obstruction. The zero-knowledge mode that is produced
on this hardware is the static-circuit Aurora run (last row;
\S\ref{ssec:aurora-ref}). \textsuperscript{$\dagger$}The zero-knowledge-enabled
Aurora run is now measured (last row); the zk-disabled row is retained for
comparison. Both Aurora rows achieve $107.5$ bits, below the $\lambda=128$
target. \textsuperscript{$\ddagger$}RISC~Zero's succinct receipt carries
zero-knowledge blinding, but the vendor disclaims a formal zero-knowledge
guarantee, so we treat the receipt as not dependably
zero-knowledge~\cite{risc0zkadvisory,l2beatzk} (Section~\ref{sec:security}).
The SHA-3 row is a control, not a deployable PLUM variant.}
\label{tab:proof-mode}
\end{table}

\subsection{A Cost Model and the Substrate Crossover}\label{ssec:costmodel}
% The STUDY spine: turn point-measurements into a validated predictive model +
% the comparative viability (crossover) analysis. Anchors reconciled per-hasher
% (Griffin vs SHA-3) -- the ~33x gap between CreGen-Griffin (4.69e9/verify) and
% PLUM-verify-SHA3 (~1.4e8) IS the field-mismatch tax, NOT a basis error.

The measurements above are point observations; to answer ``what would this cost
at a configuration we did not run, and when is the flexible substrate the
rational choice,'' we fit a cost model to them and validate it against the points
we have. This is the analysis that converts the measurements from a benchmark
table into a basis for a deployment decision. The model is fitted and validated
in \emph{execute-mode cycle} counts; carrying it to prove-mode wall-clock assumes
the additive form survives recursion, memory pressure, segment aggregation, and
the zero-knowledge wrap, any of which can introduce nonlinearity, so the
prove-mode reading is an extrapolation, not a measured fit.

\subsubsection{An additive cost model for the BDEC relations}
Both BDEC relations are, at the cycle level, dominated by repeated
signature verifications (Section~\ref{ssec:cost-decomposition}): CreGen contains
two, and ShowCre contains $k+2$ for $k$ shown credentials. Writing
$C_{\mathsf{ver}}(\lambda,h)$ for the execute-mode cycle cost of one PLUM
verification at security level $\lambda$ under hasher $h$, and $C_0$ for the
fixed per-run overhead (input parsing and witnessing; the ledger-membership,
revocation, and disclosure checks are performed by the relying party outside the
proof), the model is
\begin{align}
  C_{\mathsf{CreGen}}(\lambda,h) &\approx 2\,C_{\mathsf{ver}}(\lambda,h) + C_0,
  \label{eq:cost-cregen}\\
  C_{\mathsf{ShowCre}}(k,\lambda,h) &\approx (k+2)\,C_{\mathsf{ver}}(\lambda,h) + C_0,
  \label{eq:cost-showcre}
\end{align}
The model is deliberately linear in the verification count, because the cost
decomposition of Section~\ref{ssec:cost-decomposition} shows that everything else
in the proof is dominated by the $\Theta(k)$ algebraic-hash work inside the
verifications; the ledger-membership, revocation, and disclosure checks do not
enter the proving cost, as the relying party performs them outside the proof.

\subsubsection{Validation against the measured points}
The model is falsifiable against the two BDEC points we measured. The strongest
internal check is CreGen: Equation~\eqref{eq:cost-cregen} predicts the two
verifications should each cost $\approx C_{\mathsf{ver}}$ and together dominate
the trace. The measurement bears this out directly: the two verifications cost
$4.69\times10^{9}$ cycles \emph{each} (the two within $0.01\%$ of \emph{each
other}), their sum $9.39\times10^{9}$ accounting for $\approx99.8\%$ of the
$9.40\times10^{9}$ total
(Table~\ref{tab:bdec}); so for the Griffin hasher at $\lambda=80$,
$C_{\mathsf{ver}}(\textsf{80},\textsf{Griffin})\approx 4.69\times10^{9}$ cycles
and $C_0$ is small. We caution that this single relation validates the
\emph{linearity} and the \emph{verification-dominance}, not the absolute
constant at other $(\lambda,h)$; the $k=14$ projection below is labelled as such.

\paragraph{The hasher gap measures the tax.}
A single PLUM verification is markedly cheaper with the SHA-3 hasher control than
with the Griffin algebraic hash, an order-of-magnitude separation that is the
field-mismatch tax of Section~\ref{sec:theoretical} quantified at the
verification level. Matched-$\lambda$ figures exist on both substrates: on SP1
at $\lambda=80$ in execute mode, the software-Griffin verification costs
$7.08\times10^{9}$ cycles against $1.11\times10^{8}$ for the SHA-3 arm
(${\approx}64\times$), and on RISC~Zero at $\lambda=80$ the Griffin
per-verification cost of $4.69\times10^{9}$ cycles stands against
${\approx}1.44\times10^{8}$ cycles for the SHA-3-hasher verify
(${\approx}33\times$, field multiplications routed through
\texttt{sys\_bigint}). The qualitative conclusion (that the
algebraic hash is the tax-bearing component while the field arithmetic is not, so
$C_{\mathsf{ver}}$ must be reported per hasher) holds regardless of the exact
multiplier. The precompile narrows
this gap (Table~\ref{tab:ablation}); closing it entirely is what a native
Griffin AIR on the proving path would buy.

\paragraph{ShowCre execute-mode cost (measured at $k=1,2$).}
The ShowCre relation is realised by the guest
\texttt{bdec\_showcre\_plum\_griffin}, the $k+2$ generalisation of the CreGen
guest, and measured in execute mode at the two running examples. At $k=1$ (the
$\phi_1$ workload) ShowCre executes in $1.41\times10^{10}$ cycles, and at $k=2$
(the $\phi_2$ workload) in $1.88\times10^{10}$ cycles, each with all $k+2$
verifications accepting. Both agree with Equation~\eqref{eq:cost-showcre} at
$C_{\mathsf{ver}}(\textsf{80},\textsf{Griffin})\approx 4.69\times10^{9}$: the
per-verification cost is constant at $\approx 4.69\times10^{9}$ cycles across
CreGen and both ShowCre instances, so the additive $(k+2)\,C_{\mathsf{ver}}$ form
is measured rather than assumed. Applying the same measured constant to a stress-test disclosure-set size, well
beyond the running examples ($k=1,2$) and the $k\le 4$ design target, projects
$\approx 7.5\times10^{10}$ cycles at $k=14$. A direct $k=14$ execute run does not reach this
projection: the $16$ accumulated verifications exhaust the guest's default
non-reclaiming bump allocator before the cycle count completes, an in-guest memory
limit distinct from the prover-side provisioning wall and clearable only by the
reclaiming allocator at additional cycle cost. The $k=1$ and $k=2$ executions are
correct and their cost is measured; the zero-knowledge \emph{proof}, however,
rests on the same anonymity-providing wrap blocked by the recursion-shape
provisioning wall of \S\ref{ssec:resource-frontier}. A zero-knowledge
proof of ShowCre at any $k\ge 1$ is therefore bounded by the provisioning wall
rather than by the CreGen prove anomaly (\S\ref{ssec:cregen-anomaly}). Clearing
the provisioning wall is a structural fix, a one-time regeneration of SP1's
recursion-shape artifacts rather than added memory, and a post-quantum-anonymous proof
additionally requires a post-quantum-sound wrap (Section~\ref{sec:security}).

\paragraph{Calibrating the trace-area constants.}
Calibrating the constants of the trace-area model
(Equation~\eqref{eq:trace-area-model}) on the two completing SP1 prove runs of
Table~\ref{tab:plum-verify} (the Cell~2 Griffin arm and the Cell~3 SHA-3
control; \S\ref{ssec:headline}), the control arm prices
core work at $\kappa_{\mathsf{core}}\,\bar a_{\mathsf{core}}\approx
7.2\,\mu\mathrm{s}$ per guest cycle, and attributing the Griffin arm's residual
$17.5$~minutes to its $5.6\times10^{7}$ cells gives an effective
$\kappa_{\mathsf{Griffin}}\approx 19\,\mu\mathrm{s}$ per committed cell; in
unit-free form, committing one Griffin-chip cell costs the prover about as much
as removing $2.6$ guest cycles buys. Two measured points fit two constants, so
the magnitude is reproduced by construction; the out-of-fit content is the sign
and the orderings the same constants imply. The calibration uses the May run
pair, for which matched execute-mode guest-cycle counts exist; the June
re-measurement (\S\ref{ssec:headline}) confirms the postdicted sign, and we do
not re-derive the constants for it, lacking matched June cycle counts. Those orderings are consistent with
the rest of Table~\ref{tab:plum-verify}: the precompile-free baseline's $7.08\times10^{9}$ cycles
price at upwards of $14$~hours of core work alone (before the chip area of its
$4{,}870{,}724$ \texttt{UINT256\_MUL} events is charged), consistent with its
DNF, and against that baseline the precompile still pays by orders of magnitude,
removing about $6.6\times10^{6}$ cycles per permutation against
$5.3\times10^{4}$ cells added.

\subsubsection{The substrate crossover}\label{sssec:crossover}
The decision a deployer faces (Section~\ref{sec:discussion}) is not about which
substrate is faster per proof, since the static circuit wins that in any case. It
is about which substrate is cheaper over a deployment \emph{lifetime} during which
the predicate changes. Let the lifetime
contain $P$ proof generations and $U$ localised predicate changes. Amortised
total cost is, schematically,
\begin{align}
  \text{static:}\quad & U\cdot R_{\mathsf{static}} + P\cdot t^{P}_{\mathsf{static}},
  \label{eq:static-cost}\\
  \text{zkVM:}\quad & U\cdot 0 + P\cdot t^{P}_{\mathsf{zkVM}},
  \label{eq:zkvm-cost}
\end{align}
where $R_{\mathsf{static}}$ is the per-change recompilation cost and $t^{P}$ is
per-proof time. We isolate $R_{\mathsf{static}}$ directly (\S\ref{ssec:churn-eval}):
for the single-Loquat-verification R1CS at the $2^{19}$ BDEC building-block size
it is R1CS construction ($\approx 0.2$~s) plus Aurora
parameter setup (measured at $\approx 3\,\mu$s, because Aurora is
non-preprocessing), so $R_{\mathsf{static}}\approx 0.3$~s. The multi-minute figure
of Table~\ref{tab:aurora-ref} is the per-\emph{proof} prover, not the per-change
recompile, and does \emph{not} enter $R_{\mathsf{static}}$. The zkVM's update term is zero
(Table~\ref{tab:churn}). The static regime is cheaper only while
\[
  U\cdot R_{\mathsf{static}} < P\,\bigl(t^{P}_{\mathsf{zkVM}}-t^{P}_{\mathsf{static}}\bigr),
\]
i.e.\ while the relation-change rate $U/P$ stays below a crossover
\[
  \Bigl(\tfrac{U}{P}\Bigr)^{\!*} \;=\; \frac{t^{P}_{\mathsf{zkVM}}-t^{P}_{\mathsf{static}}}{R_{\mathsf{static}}}.
  \label{eq:crossover}
\]
The measured $R_{\mathsf{static}}\approx 0.3$~s is decisive, and it cuts
\emph{against} a naive flexibility argument. Because $R_{\mathsf{static}}$ is
sub-second while the zkVM's per-proof penalty exceeds the static prover's
($t^{P}_{\mathsf{zkVM}}>t^{P}_{\mathsf{static}}$ even granting the cross-scheme seam
of \S\ref{ssec:aurora-ref}: static Loquat-$\mathbb{F}_{p^{127}}$, zkVM
PLUM-$\mathbb{F}_{p}$), the crossover $\bigl(U/P\bigr)^{*}$ of
Eq.~\ref{eq:crossover} is on the order of thousands of relation changes per proof,
a regime no realistic deployment reaches. \textbf{Against a transparent,
non-preprocessing argument such as Aurora, the zkVM thus has no wall-clock
flexibility advantage:} both regimes rebuild a sub-second artefact per change, and
the zkVM is then dearer per proof. Recompile wall-clock does not capture the
genuine per-change cost of the static regime; that cost is the \emph{re-audit and
engineering} footprint of a new relation-specific circuit
(Table~\ref{tab:churn}, $N_{\mathsf{circ}}/N_{\mathsf{param}}$). A \emph{wall-clock} flexibility advantage would appear only
against a \emph{preprocessing} argument (Groth16~\cite{groth2016}, Marlin~\cite{chiesa2020marlin}, or circuit-specific
PLONK~\cite{gabizon2019plonk}), whose per-circuit key generation makes $R_{\mathsf{static}}$ large, and
Aurora, the BDEC baseline, is not such a system. We therefore report the
flexibility result as \emph{qualitative} (the regeneration footprint of
Table~\ref{tab:churn}) together with this measured wall-clock caveat, rather than
as a numeric crossover that favours the zkVM.

\subsection{Update Churn}\label{ssec:churn-eval}

Table~\ref{tab:churn} enumerates the regeneration footprint of
Definition~\ref{def:update-churn} for the three localised verification-predicate changes (Definition~\ref{def:lpc}) of
\S\ref{ssec:update-churn}, using $\phi_1\!\to\!\phi_2$
(par.~\ref{par:phi-examples}) as the presentation-change instance. Here
$\phi_1\!\to\!\phi_2$ is a presentation-\emph{structure} change: it moves the
credential count from $k{=}1$ to $k{=}2$, altering the relation
$R_{\mathsf{show}}^{\Sig}$ and hence the static R1CS. A presentation \emph{policy}
change at fixed structure (e.g.\ retuning a threshold the verifier checks on the
disclosed attributes) is handled verifier-side and triggers no regeneration on
either substrate, so the rows below measure structural changes.

\begin{table}[h]
\centering
\small
\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Static circuit (Aurora)} & \multicolumn{2}{c}{zkVM} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}
Localised change & $N_{\mathsf{circ}}$ & $N_{\mathsf{param}}$ & $N_{\mathsf{circ}}$ & $N_{\mathsf{param}}$ \\
\midrule
Presentation change ($\phi_1\!\to\!\phi_2$, $k{:}1{\to}2$) & $\geq 1$ & $\geq 1$ & $0$ & $0$ \\
Signature swap (Loquat $\to$ PLUM)              & $\geq 1$ & $\geq 1$ & $\geq 1$\textsuperscript{$\ast$} & $\geq 1$\textsuperscript{$\ast$} \\
Security retune ($\lambda{=}80\to128$)          & $\geq 1$ & $\geq 1$ & $0$ & $0$ \\
\bottomrule
\end{tabular}
\caption{Circuit/parameter regeneration footprint ($N_{\mathsf{circ}},N_{\mathsf{param}}$
of Definition~\ref{def:update-churn}). The static-circuit regime regenerates and
re-audits its R1CS and constraint-dependent parameters on every localised change;
the zkVM absorbs a change as a guest-program edit, leaving the universal relation
unchanged ($N_{\mathsf{circ}}=N_{\mathsf{param}}=0$), except where the change
crosses the precompile boundary. \textsuperscript{$\ast$}The Loquat~$\to$~PLUM
swap moves the hash field from $\mathbb{F}_{p^2}$ (127-bit) to the 199-bit
$\mathbb{F}_p$, so the field-specific Griffin AIR is regenerated
(\S\ref{ssec:update-churn} carve-out). The components
$N_{\mathsf{src}},N_{\mathsf{audit}}$ are reported in prose below.}
\label{tab:churn}
\end{table}

\paragraph{Structural separation, and the measured recompile cost.}
The structural separation in Table~\ref{tab:churn} is established by
construction: a zkVM's relation is fixed to the ISA, so a presentation-structure
or signature change is absorbed as a source edit
($N_{\mathsf{circ}}=N_{\mathsf{param}}=0$) with no recompilation. \emph{Both}
components of the static side are now measured, and the result overturns the
earlier conjecture that recompilation is expensive. R1CS
\emph{construction} from the BDEC relation is sub-second (${\approx}0.2$~s of
CPU time per emission at Loquat-$\lambda{=}80$ and $\lambda{=}128$, single timed
runs; cached wall-clock $0.48$~s). The Aurora
\emph{parameter setup} that a relation change forces to recompute is, measured
directly, about $3\,\mu$s (the runner reports $t_{\mathrm{setup}}=0.003$~ms at the
$2^{19}$-padded single-Loquat-verification R1CS, the BDEC building-block size),
because Aurora is \emph{non-preprocessing}: it has
no per-circuit indexer. The prover's low-degree-encoding FFTs (its dominant step) run \emph{per proof},
not per relation change, and so do not enter $R_{\mathsf{static}}$. Hence $R_{\mathsf{static}}\approx 0.3$~s at the
BDEC building-block size. A $\phi_1\!\to\!\phi_2$ change alters only the credential count and does
not enlarge the microsecond parameter setup, so this bound also holds for the
per-presentation delta; a $k$-parameterised emitter would refine the
R1CS-construction term but cannot move the dominant conclusion, namely that for a
transparent non-preprocessing argument the per-change recompile is sub-second.
Consequently the zkVM's wall-clock flexibility advantage over Aurora is negligible,
and the static regime's real per-change cost is the \emph{re-audit} of a new
circuit (Table~\ref{tab:churn}) rather than recompile time.

\paragraph{The source-edit and audit components.}
The remaining tuple components, promised in prose by
Section~\ref{ssec:update-churn}, are read off this project's own repository
record. For the security retune ($\lambda=80\to128$) the zkVM side is the
clean case: the guest binary is parameterised over $\lambda$, so
$N_{\mathsf{src}}=0$ and $N_{\mathsf{audit}}=0$; the static side keeps its
generator source ($N_{\mathsf{src}}=0$) but its regenerated matrices re-enter
the review boundary ($N_{\mathsf{audit}}\ge 1$). For the presentation change
($k{:}1\to2$) the zkVM edits one guest source (the presentation relation's
credential count), giving $N_{\mathsf{src}}=1$ and $N_{\mathsf{audit}}=1$ (the
changed guest and its image identifier); the static side regenerates the merged
R1CS, with the regenerated circuit as the re-audited component. For the
signature swap (Loquat $\to$ PLUM) the static side replaces the constraint
generator for the verification predicate and re-audits the full regenerated
circuit. The zkVM side is the honest headline of this dimension: the guest edit
itself is small, but because the swap moves the hash field, it crossed the
precompile boundary, and in this project's record the crossing cost was the
bespoke Griffin-$\mathbb{F}_p^{192}$ chip, approximately $3{,}900$ lines of
chip and executor code (\S\ref{sec:system}), with the chip and its soundness
argument (Appendix~\ref{app:proofs}) entering the audit boundary as new
components. The absorb-as-software property therefore held for every benchmark
change except the one that touched a precompiled primitive, where the zkVM
regime paid a circuit-engineering cost of the same kind it was adopted to
avoid.

\paragraph{The advantage is conditional on the baseline's preprocessing model.}
The negligible recompile cost above is a property of Aurora's \emph{non-preprocessing}
structure; it does not generalise to static circuits at large. A \emph{preprocessing} (holographic)
transparent argument such as Fractal~\cite{cryptoeprint:2019/1076} must instead build
a per-circuit index, and that index is exactly what a relation change forces to
regenerate. Timing Fractal's indexer (its \texttt{fractal\_snark\_indexer} stage) in
isolation on the same $127$-bit-field harness as \S\ref{ssec:aurora-ref}
(${\approx}107$-bit achieved, zero-knowledge disabled; a Loquat
proxy for PLUM) gives the contrast of Table~\ref{tab:rstatic-fractal}. The
$2^{10}$--$2^{14}$ rows are synthetic R1CS instances emitted on the same harness
to expose the indexer's scaling; only the $2^{19}$ row is the
single-Loquat-verification relation at the BDEC building-block size. Aurora's
parameter setup is microseconds and flat in circuit size, whereas Fractal's holographic
indexer grows superlinearly and, at the $2^{19}$-padded size, is killed under
memory pressure mid-indexer (during a $2^{27}$-element FFT) before returning.

\begin{table}[h]
\centering
\small
\begin{tabular}{lcc}
\toprule
R1CS size (constraints) & Aurora param.\ setup (one component of $R_{\mathsf{static}}$) & Fractal indexer (one component of $R_{\mathsf{static}}$) \\
\midrule
$2^{10}$ ($1{,}024$)          & $0.0021$~ms & $0.86$~s \\
$2^{12}$ ($4{,}096$)          & $0.0023$~ms & $3.74$~s \\
$2^{14}$ ($16{,}384$)         & $0.0025$~ms & $18.5$~s \\
$2^{19}$ ($524{,}288$; Loquat single-verify) & $0.0029$~ms & OOM-killed at ${\approx}33$~min \\
\bottomrule
\end{tabular}
\caption{Per-change recompile components for a non-preprocessing
(Aurora) versus a preprocessing (Fractal) transparent argument on the same
$127$-bit-field harness, M5~Pro 24~GB. The $2^{10}$--$2^{14}$ rows are synthetic
R1CS instances; the $2^{19}$ row is the single-Loquat-verification relation at
the BDEC building-block size. Aurora has no per-circuit indexer (parameter setup
flat at ${\sim}\mu$s, so $R_{\mathsf{static}}$ stays sub-second); Fractal's
indexer is superlinear and exceeds $24$~GB at the $2^{19}$ size (OOM
mid-indexer). The zkVM regenerates no circuit.}
\label{tab:rstatic-fractal}
\end{table}

\noindent The flexibility reading is therefore \emph{conditional on the static
baseline's preprocessing model}: the zkVM's wall-clock recompile advantage is
negligible against a non-preprocessing argument (Aurora) and large against a
preprocessing one (Fractal), where at the BDEC circuit size the baseline's per-circuit
indexing is itself infeasible on the target hardware. The same-scheme and field seam
of \S\ref{ssec:aurora-ref} applies throughout (a Loquat proxy, not PLUM), and a
smaller folding parameter than the Aurora-matched $\mathsf{RS}=5$ might let the Fractal
indexer fit and yield a finite but still large $R_{\mathsf{static}}$.

\subsection{What the Measurements Establish}\label{ssec:scorecard}

We close by reading the measurements back against the cost framework of
Section~\ref{sec:theoretical}.

\paragraph{Feasibility and the field-mismatch tax.}
Without the precompile a single PLUM verification does not complete on the target
hardware (out-of-memory at ${\approx}1\mathrm{m}45\mathrm{s}$,
Table~\ref{tab:plum-verify}); with it the verification completes in $32.53$~min.
At the cycle level the bespoke Griffin precompile removes a $56.4\times$ factor at
$\lambda=80$ on SP1 (\S\ref{ssec:attribution}), and the modular-multiplication precompile
(\texttt{UINT256\_MUL} / \texttt{sys\_bigint}) separately removes $37.9\%$/$70.5\%$
of SP1/RISC~Zero cycles at $\lambda=128$ on the SHA-3-hasher path
(Table~\ref{tab:ablation}). This is the field-mismatch tax
of Definition~\ref{def:fmt} made concrete: the precompile turns an infeasible
workload into a feasible one, and the residual cost is what the rest of the
evaluation measures. The with/without contrast at the prove level is an
infeasible-to-feasible transition rather than a finite ratio (\S\ref{ssec:headline}); the
factors above are cycle-level and single-verification, and we do not promote them
to a CreGen-relation prove speedup we did not run.

\paragraph{Where the cost lives.}
The cost attribution of \S\ref{ssec:attribution} bears out the decomposition of
\S\ref{ssec:cost-decomposition}: the dominant term is the algebraic hash, whose
internal field operations inflate the $\mathbb{F}_p^{192}$ multiplication count
$70\times$ when not absorbed by the bespoke chip ($4{,}870{,}724$ against
$69{,}396$ \texttt{UINT256\_MUL} syscalls at $\lambda=80$), far beyond the
explicit verifier field arithmetic.
We report this in the zkVM's own cycle units and caution against transferring
PLUM's circuit-SNARK ``$91\%$ of constraints are hash'' figure to the cycle model.

\paragraph{Per-proof cost against the static circuit.}
For a \emph{fixed} predicate the specialised static circuit is the faster prover,
as expected from a specialised encoding: Aurora proves the single Loquat-verification circuit in
$3.76$~min (Table~\ref{tab:aurora-ref}) against $32.53$~min for precompiled PLUM
verification (Table~\ref{tab:plum-verify}). The prove times do \emph{not}
converge; but the two points are not the same scheme (Loquat over
$\mathbb{F}_{p^2}$ versus PLUM over its $199$-bit prime field, \S\ref{ssec:aurora-ref}),
so we state the gap as indicative, and the comparison the thesis turns on is the
deployment-lifecycle one below rather than the per-proof one. For Loquat the
same-scheme comparison is measured (Table~\ref{tab:loquat-substrate}); for PLUM,
closing the per-proof gap requires building an $\mathbb{F}_p$
($199$-bit) Aurora harness to run PLUM statically (the current Aurora harness is
$\mathbb{F}_{p_{127}}$-only), and we identify this as the required follow-up experiment.

\paragraph{Update churn.}
The structural separation of Definition~\ref{def:update-churn} holds by
construction (Table~\ref{tab:churn}): every localised change regenerates at least
one circuit and one parameter set on the static side and none on the zkVM side,
provided the change does not alter a precompiled primitive or its field; a change
to the hash family or field width regenerates the Griffin AIR and so incurs
$N_{\mathsf{circ}}\ge1$ on the zkVM side too (Definition~\ref{def:update-churn}).
The recompile wall-clock of the static side is now measured
($R_{\mathsf{static}}\approx 0.3$~s; \S\ref{ssec:churn-eval}) and is sub-second
because Aurora is non-preprocessing; the gap that remains is a \emph{same-scheme}
per-proof static prover time for PLUM (\S\ref{ssec:aurora-ref}) rather than the
recompile, the Loquat case being settled by Table~\ref{tab:loquat-substrate}.

\paragraph{The zero-knowledge frontier.}
The base proof completes; what binds on consumer hardware is the zero-knowledge
wrap. The full account of the measured resource failures, including the
separate lattice base-cost datum, is in \S\ref{ssec:resource-frontier}. The
wrap obstruction is substrate-specific: the static circuit completes a
zero-knowledge proof of the same Loquat building block in about $22$~minutes
(mean of three runs, $21.4$--$22.9$) on the same machine (\S\ref{ssec:aurora-ref}). The frontier claim rests on
this resource-failure evidence alone; the CreGen prove anomaly
(\S\ref{ssec:cregen-anomaly}) plays no part in it.

\subsection{Threats to Validity}\label{ssec:threats}

\begin{itemize}
  \item \textbf{Single-run and repeated-run variance.} The PLUM prove-mode
  wall-clock figures (Table~\ref{tab:plum-verify}) are single runs. The Loquat
  substrate legs are now three runs each
  (\texttt{docs/measurements/pub\_hardening\_20260610}), and they characterise
  the variance directly: the Aurora leg is tight ($223.2$--$226.6$~s,
  $\pm0.8\%$), while the SP1 leg spans $44.4$--$64.0$~minutes ($1.44\times$) on
  an otherwise idle machine, the slow run starting on a package heated by the
  preceding one. A separate datum measures contention: an Aurora run under
  concurrent prover load moved from $225.5$~s to $370.0$~s ($1.64\times$;
  \texttt{docs/measurements/loquat\_substrate\_20260610/}). Thermal state and
  co-running load on laptop-class hardware therefore move wall-clock by factors
  of $1.4$--$1.6$. Conclusions resting on order-of-magnitude separations, the
  $12$--$17\times$ substrate gap and the memory-exhaustion walls, are robust to
  movement of this size. The Cell~2 versus Cell~3 inversion is now confirmed by
  $n=2$ current-build repeats per arm: the sign is robust (Griffin slower in
  every run, May and June) and the same-session June magnitude ($1.10\times$,
  the two Cell~2 runs within $0.1\%$) is tight, while the cross-build magnitude
  (up to $2.45\times$) is build-sensitive and reported as a range
  (\S\ref{ssec:headline}). The sign is model-postdicted; the magnitude we do not
  over-read.
  \item \textbf{Thermal throttling.} Runs are on a laptop-class M5~Pro; sustained
  multi-minute proving may throttle, which would inflate wall-clock relative to a
  thermally-unconstrained host.
  \item \textbf{Hardware/OS-string provenance.} The reported platform string is
  from \texttt{sysctl}; we have not cross-checked every sub-component against an
  independent inventory.
  \item \textbf{Incomplete statement binding.} The RISC~Zero prototype does not
  yet bind the receipt journal to the public statement $(c,h,\mathsf{ppk})$
  (Section~\ref{sec:system}); until it does, an accepting receipt is a functional
  benchmark and not a statement-bound credential proof, which is a security gap and not a timing one.
  \item \textbf{Unresolved CreGen prove anomaly.} The $15.6$-hour BDEC
  credential-generation prove run ended in an internal-verifier rejection
  (\S\ref{ssec:cregen-anomaly}), not a resource limit; it is a single run with
  undetermined cause, no claim of this thesis rests on it, and a
  segment-diagnostic re-run is required before any statement about it. It is
  explicitly \emph{not} part of the deployability frontier
  (\S\ref{ssec:resource-frontier}).
  \item \textbf{Parameter-set seam.} Prove-mode wall-clock is at $\lambda=80$;
  cycle attribution is at $\lambda=128$. The two are never combined into one
  figure, but a reader must track which is which; the tables are labelled
  accordingly.
  \item \textbf{Non-uniform precompile.} SP1 uses the bespoke Griffin AIR;
  RISC~Zero uses \texttt{sys\_bigint}. Cross-zkVM numbers reflect this mechanism
  difference and are not a like-for-like substrate comparison.
  \item \textbf{Constraint-level soundness of the Griffin AIR is empirically
  supported, not proved} (the chip--constraint equality, Assumption~\ref{asm:air}, of Section~\ref{sec:security}); the positive
  feasibility figures, namely the $32.53$-minute Cell~2 prove, the
  $13.0$-second smoke proof, and any prove-level reading of the $56.4\times$
  cycle ratio, assume
  the chip computes the reference permutation: proved at the per-row layer
  (Appendix~\ref{app:proofs}, conditional on the byte-range lookups under
  the lookup-binding hypothesis), tested on $256$ random vectors and by single-cell fault
  injection on seven of the ten constraint families, and not verified at the
  constraint level for the lookup-borne families or the code-to-constraint
  transcription. The negative findings of this
  chapter do not carry this premise: an unsound AIR would mean the measured
  Griffin-arm cost understates that of a corrected and at-least-as-costly chip,
  so the non-terminating precompile-free baseline, the SHA-3 inversion, the
  recursion-shape provisioning wall, and the Loquat substrate gap survive unchanged or
  strengthen, as argued in Section~\ref{sec:security}.
  \item \textbf{Control vs.\ deployable.} The SHA-3 arm is a diagnostic control;
  it is not a secure PLUM variant and its timing is not a deployment claim.
\end{itemize}

\subsection{Summary}\label{ssec:eval-summary}
On consumer hardware, the precompile suite moves PLUM verification from infeasible
to feasible (Table~\ref{tab:plum-verify}), with the cost dominated by
large-field/algebraic-hash arithmetic and reduced by $37.9$--$70.5\%$ at the cycle
level (Table~\ref{tab:ablation}); the full BDEC credential relation's honest path
is validated in execute mode, while its succinct prove ended in an internal-verifier
rejection categorically distinct from the deployability frontier (Table~\ref{tab:bdec},
\S\ref{ssec:cregen-anomaly}), the binding consumer-hardware frontier being the
zero-knowledge wrap (\S\ref{ssec:resource-frontier}); and the zkVM regime is free of
circuit and parameter regeneration for every localised change that leaves the
precompiled primitive suite intact (hash-family or field-width changes regenerate
the Griffin AIR; \S\ref{ssec:churn-eval}, Table~\ref{tab:churn}). The one comparison the data does not yet settle is a \emph{same-scheme} per-proof prover time for PLUM; for Loquat it is settled in the static circuit's favour (Table~\ref{tab:loquat-substrate}), and the static-circuit recompile wall-clock is now measured ($R_{\mathsf{static}}\approx 0.3$~s) and is not the bottleneck. The next chapter takes up what these results mean for the deployment decision and which questions they leave open.
