\section{Conclusion}\label{sec:conclusion}

This thesis asked whether a general-purpose zero-knowledge virtual machine (zkVM) is a viable substrate for post-quantum anonymous credentials on consumer hardware. The goal was to make BDEC credential verification, instantiated with the PLUM signature, generate proofs in a practically acceptable time inside a zkVM on a personal computer (Apple M5~Pro, $24$~GB); the means was a custom precompile suite that absorbs the field-mismatch tax of PLUM's large-prime-field algebraic hashing. We realised the construction on RISC~Zero and SP1, measured it against an Aurora static-circuit baseline, and analysed what the substitution preserves and what it costs.

\subsection{Main Findings}\label{subsec:conclusion-findings}

The central reusable result is the cost criterion of Section~\ref{sec:theoretical}: a precompile lowers core-proving cost only on the field-mismatch axis, and whether it pays is a trace-area inequality whose sign the measured hash-cost inversion confirms (the Griffin precompile proves slower than a SHA-3 control, $14.25$ versus $13.05$~min). The principal \emph{deployability} finding, secondary to the criterion above, is a two-pronged obstruction to post-quantum anonymity. On the target hardware the proof one can afford to generate (the standalone PLUM-verification receipt) is not zero-knowledge, and the zero-knowledge wrap the toolchain exposes is not post-quantum (Section~\ref{ssec:zk-gap}). The first prong is structural rather than a memory bound: at default prover concurrency the PLONK wrap of the standalone PLUM verification receipt is killed by memory pressure at ${\approx}20$~GB of the $24$~GB budget, but throttling the concurrency runs it well within budget, where it fails instead at recursive compression, the recursion-shape provisioning wall: the toolchain's static recursion shapes are provisioned for smaller relations than PLUM verification, a one-time regeneration away from completing; each stage of the wrap fits within $24$~GB (Section~\ref{sec:evaluation}), so the binding cost is the regeneration's compute, with the regenerated toolchain not run end-to-end. The second is a fact about the deployed toolchains, and it takes a different form on each substrate: the only succinct zero-knowledge wraps exposed by the SP1 pipeline we measure (v6.2.1, forked) are Groth16 and PLONK, both pairing-based, so on SP1 even a completed wrap could be at best classically secure; on RISC~Zero the blinded succinct receipt carries no formally established zero-knowledge guarantee, so BDEC's anonymity premise cannot be formally discharged on it, and its pairing-based wrap is x86-only and does not run on the Apple-silicon target at all (Section~\ref{sec:security}). This obstruction on both sides holds independently of any timing number and of the open AIR-soundness Assumption~\ref{asm:air} (the dependency direction is recorded in Section~\ref{sec:security} and restated under Limitations below).

Around that finding, the measurements establish the surrounding cost picture. The custom Griffin-$\mathbb{F}_p^{192}$ precompile turns a non-terminating emulation into a finite measurement: without it, standalone PLUM verification exhausts memory at ${\approx}1$m$45$s on the target hardware; with it, the verification proves in $14.25$ minutes (mean of three fixed-config runs, $14.05$--$14.61$; Section~\ref{sec:evaluation}) on SP1 at $\lambda=80$, under the documented reduced-memory configuration, default settings exhausting memory (Section~\ref{sec:evaluation}). Two of the measurements falsify natural predictions. First, the bespoke algebraic-hash precompile is \emph{slower} than a software SHA-3 control at matched security in every run we took, on both the May and the June builds. The sign is robust; under fixed-config replication ($n=3$ per arm) the magnitude is ${\approx}1.09\times$ in prove wall-clock ($14.25$ versus $13.05$ minutes). Arithmetising a $199$-bit-field algebraic hash through a custom chip costs more than a hash the substrate already supports. The trace-area accounting of Section~\ref{sec:theoretical} postdicts the sign of this inversion from the chips' committed dimensions: the custom chip adds roughly $5.6\times10^{7}$ committed trace cells per verification (a count derived from the chip's committed dimensions, not a logged trace readout) that the control never pays, a term the guest-cycle accounting alone does not carry. Second, against the transparent, non-preprocessing Aurora baseline the zkVM's update-churn advantage never shows up in wall-clock. The static-side recompile is sub-second ($R_{\mathsf{static}}\approx 0.3$~s, Section~\ref{ssec:churn-eval}), so what separates the two regimes is the regeneration-and-re-audit \emph{footprint} (Table~\ref{tab:churn}), a cost in engineering and assurance rather than in proving time. A wall-clock flexibility advantage would appear only against a \emph{preprocessing} baseline, which is not what BDEC uses.

The credential relations themselves run correctly inside the zkVM: credential generation executes both of its signature verifications to acceptance in $9.4\times10^{9}$ guest cycles, and the presentation relation in $1.41\times10^{10}$ (one credential shown) and $1.88\times10^{10}$ (two) cycles, consistent with the additive cost model. On SP1 both relations complete a succinct, non-zero-knowledge prove (credential generation in $26.98$ and $31.11$~min, presentation in $40.15$ and $44.24$~min at $k{=}1$ and $58.14$~min at $k{=}2$; Section~\ref{sec:evaluation}). No zero-knowledge \emph{proof} of either relation was produced on consumer hardware: on RISC~Zero the credential-generation succinct prove terminated in an internal-verifier rejection (RISC~Zero~$3.0.5$; an anomaly under diagnosis, categorically distinct from a resource limit and on which no claim of this thesis rests, Section~\ref{ssec:cregen-anomaly}), and the presentation wrap is blocked by the same recursion-shape provisioning wall.

\subsection{Answer to the Research Question}\label{subsec:conclusion-answer}

The answer is \emph{not yet}, for a reason more specific than ``too slow.'' Standalone PLUM verification becomes feasible on consumer hardware, at $\lambda=80$ on SP1 under tuned memory settings and conditional on Assumption~\ref{asm:air}, once the precompile absorbs the field-mismatch tax. Deployable post-quantum \emph{anonymity} is not reached, because it requires a zero-knowledge wrap that is at once feasible to generate within $24$~GB and post-quantum-sound, and neither is provided by the pipelines measured here, for the substrate-specific reasons of the previous subsection. The substrate choice is therefore governed by the deployment-lifecycle decision rule of Section~\ref{sec:discussion}: where the verification predicate is fixed and proofs are frequent, the specialised static circuit is preferable; where the predicate, signature, or security parameters evolve over the deployment's lifetime, the zkVM's fixed universal relation confines each change to a guest-program edit, provided the change leaves the precompiled primitive suite intact, since a hash-family or field-width change regenerates the Griffin AIR. Against a non-preprocessing baseline its advantage there is a smaller re-audit footprint, not a faster runtime.

\subsection{Limitations}\label{subsec:conclusion-limitations}

The results are bounded in ways we state explicitly. Prove-mode wall-clock is measured at $\lambda=80$, the largest level at which it is observable within the memory budget and not a deployment security level; the field-multiplication ablation is reported at $\lambda=128$, while the Griffin-precompile cycle ratio ($56.4\times$) and the credential relation cycle counts are execute-mode figures at $\lambda=80$. The static baseline is Loquat over $\mathbb{F}_{p_{127}}$ rather than same-scheme PLUM, with the per-proof figure measured at zero knowledge disabled and the zero-knowledge-enabled counterpart proving in about $22$ minutes (mean of three runs, $21.4$--$22.9$) (Section~\ref{sec:evaluation}), so the per-proof comparison is indicative and the lifecycle crossover is bounded in form but not located in wall-clock. The standalone PLUM Cell~2 and Cell~3 prove figures are means of three fixed-config runs (ranges reported); the credential-relation prove figures are single runs, and the Loquat substrate comparison is three runs per leg (Section~\ref{sec:evaluation}). The security guarantees are conditional: the constraint-level soundness of the Griffin AIR (Assumption~\ref{asm:air}), the existence of a PLUM signature-transcript simulator (Assumption~\ref{asm:szk}), the quantum-random-oracle lift (Assumption~\ref{asm:qrom}), and the canonical prime instantiation are carried as open assumptions (Section~\ref{sec:security}). Of these, Assumption~\ref{asm:air} conditions only the positive results, the $14.25$-minute feasibility figure and the security-preservation theorem; the negative findings, including both prongs of the obstruction, the SHA-3 inversion, the precompile-free baseline's memory exhaustion, and the absence of a wall-clock flexibility advantage, would survive or strengthen under an unsound AIR, since the measured Griffin-arm cost would then understate that of a corrected and at-least-as-costly chip (Section~\ref{sec:security}). Finally, the RISC~Zero prototype does not yet bind the receipt journal to the public statement, so the artifact is at present a functional benchmark rather than a statement-bound credential verifier. The wrap's $24$~GB-sufficiency is established at the stage level, each stage fitting the budget (Section~\ref{sec:evaluation}), not by a completed regenerated-toolchain run, and the gnark wrap figure is a program-independent probe at $n=1$.
\subsection{Future Work}\label{subsec:conclusion-future}

With the Loquat same-scheme substrate comparison now measured (Table~\ref{tab:loquat-substrate}), the most informative next measurement is the \emph{same-scheme} per-proof comparison for PLUM itself: building a $199$-bit $\mathbb{F}_p$ Aurora harness to run PLUM statically, which would test whether the measured $12$--$17\times$ Loquat substrate gap transfers to PLUM and locate the crossover that the decision rule currently bounds only in form. The open assumptions are the substantive cryptographic work that remains: Assumption~\ref{asm:air}, whose per-row layer now carries a written proof (Appendix~\ref{app:proofs}) and rejection-direction fault-injection coverage for seven of the ten constraint families, by extending the suite through a full-prover harness to the lookup-borne families (round count, memory binding, cross-row threading) and by machine-checking the code-to-constraint transcription, for instance by the method of Arguzz~\cite{hochrainer2025arguzz}; Assumption~\ref{asm:szk} by constructing a PLUM signature-transcript simulator; and the post-quantum-anonymity goal by a post-quantum-sound zero-knowledge wrap (a STARK- or lattice-based recursive argument) in place of the pairing-based wraps the toolchains currently expose. The immediate engineering step is to commit the public statement to the receipt journal, after which the credential-generation prove anomaly should be re-run with segment-level diagnostics. Confirming the canonical prime with PLUM's authors would discharge the prime-instantiation premise.

The post-quantum-anonymity obstruction of Section~\ref{ssec:zk-gap} is, by that analysis, an engineering gap rather than a fundamental one, and the path through it can be stated precisely. The zero-knowledge the produced receipt lacks can be supplied without a pairing-based wrap, by masking the prover's own low-degree test rather than wrapping its output: for the {FRI}/{STARK}-based prover of RISC~Zero by masked {FRI}~\cite{cryptoeprint:2024/1037}, and for the multilinear Basefold-style commitment of the SP1 prover by the VEIL compiler~\cite{cryptoeprint:2026/683}, both plausibly post-quantum and neither reintroducing a quantum-vulnerable assumption. Both exist as code (VEIL ships in the SP1 fork, as yet unwired); what remains is to wire one into the prove path and to measure whether the masked prover stays within the $24$~GB budget, the one quantity this thesis leaves unsettled. Closing the obstruction this way would deliver \emph{computational} post-quantum anonymity in the quantum random-oracle model. It would not, on its own, deliver \emph{everlasting} anonymity, which additionally requires the standard-model statistically-hiding trace commitment of Proposition~\ref{prop:everlasting} and remains a separate open problem.

A further direction concerns the precompile construction itself. The three chips of Section~\ref{sec:system} instantiate one interface by hand, and the cost criterion decides, per primitive, which of them is worth building. The natural generalisation is a generator that emits a sound precompile directly from a primitive's specification: given the S-box degree, state width, round structure, and field of a large-prime algebraic primitive, it would produce the AIR together with a machine-checkable soundness obligation from a single meta-theorem, rather than writing and auditing each chip separately. This would turn the interface of Section~\ref{ssec:soundness} from a contract instantiated three times into a reusable construction method, and it would let the criterion drive an automated build-or-skip decision in the manner of the autoprecompile line~\cite{powdr2025autoprecompiles}, but over native large-field arithmetic rather than merged emulation. The soundness meta-theorem is the hard part, and we leave it as future work.

\subsection{Closing}\label{subsec:conclusion-closing}

Long-lived post-quantum identity systems require both cryptographic soundness and operational evolvability, and this thesis measured the price of pursuing both on the same substrate. What it contributes is a mapped frontier. The precompile makes PLUM signature verification at the $80$-bit level finite to prove on a personal computer. The deployment-lifecycle separation between the static and zkVM regimes is real, though it surfaces as a re-audit footprint and stays out of the wall-clock numbers. The obstruction to deployable post-quantum anonymity has two prongs that survive together: the proof one can afford is feasible but not anonymous, and the anonymous wrap is not post-quantum, and no amount of additional memory, favourable timing, or resolution of Assumption~\ref{asm:air} removes either. That obstruction, named precisely and surrounded by its cost picture, is what we hand to the design of the next such system.
